Self Referencing Sequences

### 2014-11-05 draft.tex
finished the determinism section

Here, \emph{maximal length} is a local property that means that the the block cannot be extended to include the numbers preceding or succeeding the block, as they will not be the repeated number. Thus, if a sequence $\{s_i\}$ is
\begin{align*}
1,3,3,3,1,1,1,3\dots \,
\end{align*}
the first block is $B_1 = \{1\}$, the second is $B_2 = \{3,3,3\}$, the third is $B_3 = \{1,1,1\}$, and so on.

It is obvious that a sequence can be specified by giving the numbers of the sequence individually, or by specifying all the blocks $B_i$. This is enough to make the following claim.

\begin{proposition}
A self-referencing sequence $s$ is uniquely specified by the generating cycle $C$, and the starting number $a \in C$. In other words, such an $s$ is guaranteed to exist and is unique.
\end{proposition}
\begin{proof} 
The proof is straightforward. The contents of block $B_i$ are determined by the value of $i$ and $s_i$. The value of $i$ determines which number is repeated throughout the block. 
\end{proof}
%
%\begin{lemma}
%Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
%\end{lemma}
%
%\begin{proof}
%Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.
%
%\begin{figure}{hb}
%\label{fig:unequalsequences}
%\center
%\includegraphics[width=0.5\linewidth]{lemma1.pdf}
%\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
%\end{figure}
%
%This implies that every block before the $k$th block just contains 1's since each block refered it itself. This means that either $j = 2$, or the sequence is not self-referential. But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle. 
%\end{proof}
%
%
%
%We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element. Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$. For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
%\begin{quote}
%\textbf{Algorithm:}\\
%Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
%
%This procedure continues deterministically so long as there is an element to read. Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
%\end{quote}

>>>>

Here, \emph{maximal length} is a local property that means that the block cannot be extended to include the numbers preceding or succeeding the block, as they will not be the repeated number. Thus, if a sequence $\{a_i\}$ is
\begin{align*}
1,3,3,3,1,1,1,3\dots \,
\end{align*}
the first block is $B_1 = \{1\}$, the second is $B_2 = \{3,3,3\}$, the third is $B_3 = \{1,1,1\}$, and so on.

Clearly, a sequence can be specified by giving the numbers of the sequence individually, or by specifying all the blocks $B_i$. This is enough to make the following claim:

\begin{proposition}
A self-referencing sequence $\{a_i\}$ is uniquely specified by the generating cycle $C$ and the starting number $a \in C$. In other words, such an $\{a_i\}$ is guaranteed to exist and is unique.
\end{proposition}
\begin{proof} 
The proof is constructive. The self-referential property of the sequence implies that the contents of block $B_i$ are entirely determined by the values of $a$, $i$, and $s_i$. The value of $i$ determines which number is repeated throughout the block. For example, when $i = 1$, the number is the starting number $a$, and when $i = 2$, we use the number following the starting number in $C$, and so on for larger values of $i$. The value of $s_i$ simply dictates the length of block $B_i$.

Thus, the sequence is uniquely given by first ``writing down" $B_1$, followed by $B_2$, followed by $B_3$, etc. One detail is that it remains to check that $s_i$ is known when $B_i$ is written down. However, writing down a block always involves writing down at least one element of the sequence, so the value of $s_i$ is always known when $B_i$ is written down. Note that there is an edge case where $s_i$ is not yet written down if the starting number is 1 and $i =2$. However, in this case, the value of $s_i$ is still known (as it is given by the rule) and thus $B_i$ can still be written down.
\end{proof}
%
%\begin{lemma}
%Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
%\end{lemma}
%
%\begin{proof}
%Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.
%
%\begin{figure}{hb}
%\label{fig:unequalsequences}
%\center
%\includegraphics[width=0.5\linewidth]{lemma1.pdf}
%\caption{Self-referencing sequences $a$ and $b$ which differ first at the $j$th element.}
%\end{figure}
%
%This implies that every block before the $k$th block just contains 1's since each block refered it itself. This means that either $j = 2$, or the sequence is not self-referential. But if $j=2$ and $a_1 = b_1 = 1$, then $a_2 \neq b_2$, which means both $a_2$ and $b_2$ comes after $1$ in the generating cycle. 
%\end{proof}
%
%
%
%We present an algorithm that produces the unique self-referencing sequence given a generating cycle and a starting element. Assume the generating cycle is $(a_1 a_2 \dots a_m)$ and that the starting element is some $a_i$. For notational simplicity, if we ever write down $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
%\begin{quote}
%\textbf{Algorithm:}\\
%Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\
%
%This procedure continues deterministically so long as there is an element to read. Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
%\end{quote}
